Initial Explorations

### 2019-02-22 background.Rmd
updates

####\  
##Context & Prior Work
From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, our program trains graduate TAs to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing. We ask them to spend a shorter amount of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and the comments given should encourage student reflection. As students' writing improves, TAs can focus on smaller issues. Our bins-based report grading protocol reinforces this strategy.

Our coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable. What we need is a more direct measure of effort.

In 2018, we began trying to analyze TAs' comments from student papers directly. We extracted and tabulated all comments from approximately 500 pairs of reports. After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.

@@

* **Subject**: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
* **Structure**: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?

Raw counts for each subject/structure pair were then converted to fractional frequencies. The final frequency table (excerpted below) provided us with a direct estimate of TA effort and emphasis.

@@

##Central Problem
Analyzing TA comments this way is informative but coding by hand is unsustainable. Developing the codebook then rating the initial set of 11,000 comments required nearly 80 hours of investigator effort. Even with the codebook in hand, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.

Given 10-12,000 comments are generated EACH SEMESTER, an automated classification method is needed.

@@

##Work Plan
The work plan for this project is:

1.  Compile initial dataset and de-identify both TAs and students by replacing names with anonymous IDs.
2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
3.  Explore comment data structure, n-gram frequencies, etc., to identify potentially useful elements for feature engineering.
4.  Write an initial Naive Bayes workflow for supervised text classification.
5.  Assign TA comments to categories using the NB classifier, then compare outcomes from automated classification to hand-coded results.
6.  Using features (from Step #3) and comparison data from initial classification protocol (from Step #4) as guides, test iterations and optimize the classifier method. 
7.  Apply the optimized classifier to the original comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier.
10.  Validate this vertical Ensemble Classifier by classifying a subset of ~2000 new TA comments extracted from a non-overlapping set of student lab reports. Compare automated results for new data set to hand-coded categories.

>>>>

####\  
##Context & Prior Work
From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing. Ideally:

*  TAs spend a shorter amount of time writing fewer comments that focus on broader issues.
*  TAs provide no more than 3-5 comments per text page.
*  Comments encourage student reflection and self-correction rather than prescribe specific changes. 
*  As students' writing improves, TAs shift their focus to the next most-significant errors or issues.

Our bins-based report grading protocol reinforces this feedback strategy.

This holistic, coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy. We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable. What is needed is a more direct measure of TA effort.

In 2018, we began trying to analyze TAs' comments on student papers directly. We extracted and tabulated all comments from ~500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester. After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.

@@

* __Subject__: did the comment focus mainly on **Basic Criteria** (minimum pass/fail requirements), **Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?

Raw counts of comments for each Subject/Structure pair were then converted to fractional frequencies. The final frequency table (excerpted below) provided us with direct estimates of the most common types of comments TAs made on student work, and of which elements of writing TAs emphasize most often.

@@

##The Challenge
First, analyzing TA comments this way is informative but coding by hand is unsustainable. Developing the codebook then rating the initial set of 11,000 comments required nearly 80 hours of investigator effort. Given 10-12,000 comments are generated EACH SEMESTER, a faster classification method is essential for this NSF project to succeed.

Second, even with the codebook in hand, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" (changes in how code features are interpreted by a single rater over time). Automated classification would eliminate coder training and minimize potential for coding drift.

@@

### 2019-02-22 dataimport.Rmd
updates

---
title: "Data Import and Initial Exploration"
author: "Dan Johnson"
date: "12/17/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```
#Overview
This page has three sections. **Part 1** describes the target dataset for this project. **Part 2** outlines the data pre-checks and import process. **Part 3** describes post-import validation developed for the target dataset, and how selected elements within the dataset are recoded for analysis.

Initial visual exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html

These linked pages summarize work to date using categorical Naive Bayes as the text classification method. They provided a baseline from which I am evaluating other approaches.

*  Baseline Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2b.html
*  Randomization Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2r.html
*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayesV2yn.html
*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayes_Smry.html

Other pages provide the theoretical background for the selected methods. Each deeper analysis method is described in detail on its own separate page. These are listed on the **[LINK] Project Outline** page.

@@

#Part 1. Initial Dataset
The development dataset is an Excel table of 10,808 rows. One column has "TA comments" extracted from student lab reports in Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

@@

#Part 2. Data Pre-Checks
Prior to import, review the data table **in Excel**; check that:

@@

Most in-depth analyses will have a similar extraction and post-processing code block. In practice, it has proven easier to modify the post-processing code than to maintain multiple separate copies of the original CSV dataset for the different analyses.

>>>>

---
title: "Data Import & Checking"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

#Overview
This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.

Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html

Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html

@@

#Describe an Element

#Describe an Element

#Overview
This page describes the target dataset, data pre-checks and import process, post-import validation, and how selected elements are recoded for analysis.

Preliminary exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html

Rationale and initial steps in development of the Classifier are described here: https://adanieljohnson.github.io/default_website/bayes_history.html

Code for the current version of the Naive Bayes Classifier is located here:
https://adanieljohnson.github.io/default_website/NB_2r2c.html

#Target Dataset
The development dataset is an Excel table of 10,808 rows. One column has "TA comments" extracted from student lab reports in Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

@@

#Pre-Check and Import
Prior to import, review the data table **in Excel** and check that:

@@

Most analyses will use a similar extraction and post-processing code block. In practice, it is easier to modify the post-processing code than to maintain multiple separate copies of the original CSV dataset for different analyses.

### 2019-02-22 datastructure.Rmd
updates

---
title: "Data Import & Checking"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

@@

#Overview
This page describes the target dataset, data pre-checks and import process, post-import validation, and how selected elements are recoded for analysis.

Preliminary exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html

@@

#Target Dataset
The development dataset is an Excel table of 10,808 rows. One column has "TA comments" extracted from student lab reports in Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.

@@

@@

##Initial Library and Data Import
```{r Libraries_1}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

```{r}
#This block reads in full CSV into a starting dataframe named "base_data".
#Subsequent analyses are performed on extracted subsets of "base_data".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

If dataset imports correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.

#Part 3. Post-Import Checks
Given some data points may be entered incorrectly by instructors or students, these post-import data checks help ensure data have been properly coded and columns contain valid entries only. (This process can be automated using the ```janitor``` data cleaning package.)

```{r}
#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
course<-unique(base_data$course)

#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
ta<-unique(base_data$ta)

#"Lab" should only have entries that match the allowed limited keywords in codebook.
lab<-unique(base_data$lab)

#It is impractical to validate all randomized identifiers. Instead look at the NUMBER of unique anonymous student IDs. Quantity should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
student<-unique(base_data$student)

#Check that the list of "code.subject" topics here matches allowed terms in codebook.
subject<-unique(base_data$code.subject)

#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
structure<-unique(base_data$code.structure)
```
####\

##Post-Import Processing
Comment labels used in the project codebook contain punctuation and spacing that complicates subsequent analysis. Rather than modify the project codebook or external processes, it is simpler to recode the problematic labels after import. For example, the following code extracts a subset of rows from the full dataset, reduces those to 3 columns, and renames the **comment.subject** column as **subject**. Four problematic code labels in the subject column are replaced with simpler terms.

```{r}
#Read in TA comments from CSV file.
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

#Select rows representing the sub-groups to compare. 
comments_subset <- filter(base_data,code.subject=="1. Basic Criteria"|code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")

#Reduce larger dataframe to 3 required columns of data (subject, structure, and text of comment), and put columns in order needed.
comments_raw <- comments_subset %>% select(23,24,22)

#Rename the columns.
names(comments_raw)[1] <- "subject"
names(comments_raw)[2] <- "structure"
names(comments_raw)[3] <- "text"

#Simplify coding terms used in "subject" column.
comments_raw[,1] <- ifelse(comments_raw[,1] == "1. Basic Criteria","1_basic", ifelse(comments_raw[,1] == "2. Writing Quality","2_writing", ifelse(comments_raw[,1] == "3. Technical and Scientific","3_technical", ifelse(comments_raw[,1] == "4. Logic and Thinking","4_logic",99))))

#Change "subject" element from character to a factor for analysis.
comments_raw$subject <- factor(comments_raw$subject)
str(comments_raw$subject)
table(comments_raw$subject)
```

Most analyses will use a similar extraction and post-processing code block. In practice, it is easier to modify the post-processing code than to maintain multiple separate copies of the original CSV dataset for different analyses.

>>>>

---
title: "Data Structure"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

#Overview
This page describes how the target dataset is structured.

How data are imported then validated for analysis is described here: https://adanieljohnson.github.io/default_website/dataimport.html

@@

#Target Dataset
The dataset being used to develop this project is an Excel table containing 10,808 TA comments extracted from student lab reports submitted during Spring 2018. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

@@

Data were anonymized in Excel, replacing names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.

@@

### 2019-02-22 index.Rmd
updates

In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester. Using an iteratively generated codebook, each of the comments was separately classified in two categories:

* __Subject__: did the comment focus mainly on **Basic Criteria, Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?

#### \   
Counts of coded TA comments were tabulated according to subject and structure, and frequencies for each subject/structure pair calculated as a fraction of all comments. This 2-way contingency table provided us with a direct estimate of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.

#### \   
##Specific Goals
Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester. There also is a significant risk of "coding drift" over time, reducing accuracy.

@@

>>>>

Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester. There also is a significant risk of "coding drift" over time, reducing accuracy.

@@

#### \   
##Work Plan
The work plan for this project is:

1.  Compile an initial dataset for testing, and de-identify both TAs and students by replacing names with anonymous IDs.
2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
3.  Use basic text analysis methods to explore comment data structure, n-gram frequencies, etc., and identify potentially useful elements for feature engineering.
4.  Write a baseline workflow using Naive Bayes for supervised text classification.
5.  Assign TA comments to categories using the NB classifier, then establish baseline accuracy values using the optimized NB classifier by comparing outcomes of automated classification to hand-coded results.
6.  Using features (from Step #3) and comparison data from NB classification protocol (from Step #5) as guides, further optimize the NB classifier method. 
7.  When accuracy stabilizes, apply the optimized NB classifier to the original FULL comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier to create a vertical Ensemble Classifier.
10.  Validate the Ensemble Classifier using a subset of ~2000 new hand-coded TA comments extracted from a non-overlapping set of student lab reports. Compare automated results for new data set to hand-coded categories.

Looking beyond the current project, I will be using the NB-based Ensemble Classifier as the control method for evaluating other text classifier strategies.

### 2019-02-22 initial_exploration.Rmd
updates

>>>>

#Overview
This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.

Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html

Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html

#Describe an Element

#Describe an Element
